Detection of slow slip events using wavelet analysis of GNSS recordings

### 2021-10-10 paper/agujournaltemplate.tex
Send paper to advisors

\title{=enter title here=}

@@

\authors{=list all authors here=}

@@

\affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

@@

\correspondingauthor{=name=}{=email address=}

@@

\begin{abstract}
enter abstract here

\end{abstract}

@@

In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data. For instance, ~\citet{AGU_2009} computed the GPS-estimated moment release for 23 ETS events in Cascadia between 1997 and 2008. Simultaneously, they computed the cumulative number of hours of tectonic tremor recorded for each event. They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8. For all these events, at least 50 hours of tectonic tremor where observed simultaneously with the GPS deformation. However, many smaller bursts of tremor of duration 1 to 50 hours were also observed in between the big ETS events. Based on the relationship between slow slip moment and number of hours of tremor, ~\citet{AGU_2009} suggested that smaller slow slip events of magnitude 5-6 may occur simultaneously with the tremor bursts without being detectable in the GPS data. \\

~\citet{FRA_2016} transformed the GPS time series into daily increments of surface motion by computing the first order differentiation of the time series. He then discarded the daily increments observed during known big slow slip events, and focused on the inter-events period. He divided the daily increments into two groups: the first group contains days when slow seismicity (tectonic tremor and LFEs) is detected, the second group contains days when the numbers of LFEs (for Guerrero) or tremor (for Cascadia) is lower than a threshold. He then stacked separately the two groups of daily increments and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion. He also observed a cumulative displacement in the southern direction (for Guerrero) and the western direction (for Cascadia) corresponding to the release period when tremor and LFEs are observed. This reverse displacement corresponds to smaller slow slip events not directly observable in the GPS data.  \\

However, in other subduction zones such as New Zealand, there is no clear relationship between tremor and slow slip occurrence and these methods cannot be applied. We thus need other methods to be able to better detect and quantify slow slip. \\

Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next. Wavelet methods have been widely used for geophysical applications ~\citep{KUM_1997}. However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events. \\

~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011. The noise in the tide gauges data is associated with tides, and ocean and atmospheric noise on multiple timescales (a few days for storms to decades for oscillations between ocean basins), and is assumed to be coherent between each of the four tidal gauges studied. On the contrary, the uplift due to ETS events should be different at each tidal gauge. They first removed the tides using NOAA hourly harmonic tidal predictions. They then removed the residual noise using a method based on the DWT. More precisely, the authors applied a DWT to each of the four sites studied, and to the average of the four sites. Then, for each level of the DWT decomposition, they carried out a linear regression between the detail for one site and the detail for the average of the four sites. This process gives a coefficient for each level and for each site. They then constructed a noise signal for each site by multiplying the coefficient from the linear regression by the detail of the average over the four sites, and summing for all levels. The noise signal thus obtained was then removed from the time series. They then stacked multiple events to obtain an average event uplift rate, aligning the 12 ETS events using exact timing from GPS data. A difference in uplift between the two tidal gauges at Port Angeles and Port Townsend was then clearly seen in the stacked time series. Finally, the authors removed the long-term uplift rate and the long-term sea level rise to obtain an average inter-event uplift rate. They found that the inter-event deformation at a site is equal and opposite to the deformation during an ETS event, suggesting that ETS events are, on average, releasing the strain accumulated between ETS events. \\

~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005. They stabilized the GPS time series using a reference set of stations from stable North America. They then modelled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, and Heaviside step functions corresponding to earthquakes and hardware upgrades. The linear system was then solved using a weighted QR decomposition. Finally, they applied a Gaussian wavelet transform to the residual time series to get the exact timing of the slow slip at each GPS station. The succeeding wavelet basis functions are increasingly sensitive to temporal localization of a given signal, and the onset of faulting appears on the wavelet spectrum as an amplitude spike present over all frequencies. The offset for each slow slip event was then used to invert for the slow slip at depth by assuming a thrust fault slip at each subfault of the plate boundary. An equivalent moment magnitude was thus obtained. \\

Finally, instead of using wavelets in the time domain, ~\citet{OHT_2010} used 2D wavelet functions in the spatial domain to detect slow slip events. They designed the Network Stain Filter (NSF) to detect transient deformation signals from large-scale geodetic arrays. Contrary to their previous work on the Network Inversion Filter (NIF), there is no need to specify potential sources of deformation. They modeled the position of the GPS station by the sum of the secular velocity, a spatially coherent field, site-specific noise, reference frame errors, and observation errors. The spatial displacement field is modeled by the sum of basis wavelets (the Deslauriers-Dubuc wavelet of degree 3) with time-varying weights. The transient is considered to be nearly steady-sate, so that it has spatial weights for the displacement and the velocity, but the acceleration is modeled by a random walk with a time-varying variance. All the time varying coefficients are estimated using Kalman filtering, and the optimization problem is regularized with the spatial sum of the transient strain rate field. Their method has been successfully used to detect a transient event in the Boso peninsula, Japan, and a slow slip event in the Alaska subduction zone ~\citep{WEI_2012}. \\

In this study, we use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia. Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may be currently undetected with standard methods. \\

@@

We focused our study on northwest Washington State. For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University. These are network solutions in ITRF2008 with phase ambiguities resolved. Solutions are computed with JPL/NASA orbits and satellite clocks. North, East, and Vertical directions are available. However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio. The wavelet method works best with data with zero mean, and no sharp discontinuities, so we use the cleaned dataset, that is GPS times eries with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}. For the seismic data, we used the tremor catalog from the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}. Tremor were detected and located using waveform envelope correlation and clustering and a centroid location is available for every given five-minute time window when tremor was detected. As the catalog starts in August 2009, we only looked at GPS data recorded in 2009 or later. \\

@@

The MODWT of a time series can be defined for any length $N$. The time step of the wavelet vectors $\widetilde{W}_j$ and the scaling vector $\widetilde{V}_J$ is equal to the time step of the original time series. When we circularly shift the time series, the corresponding wavelet vectors, scaling vector, details and smooths are shifted by the same amount. The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series. The wavelet methods for time series analysis are explained in a more detailed way in ~\citet{PER_2000}). \\

@@

To increase the signal-to-noise ratio and be able to better detect slow slip events, we stack the signal over several GPS stations. We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal. We choose 16 points located on the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4). Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations. We thus have a stacked detail for each level 1 to 8 of the wavelet decomposition. \\

@@

We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. The result is shown in the top panel of Figure 6, where each line represents one of the locations. To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold. To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6. We used the opposite of the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks. Although the latitudinal extension of the events is not always the same for the GPS data and for the tremor data, we identify the same 10 events in both 8th wavelet decompositions for the 8th level: Summer 2010, Summer 2011, Summer 2012, Fall 2013, Summer-Fall 2014, Winter 2015-2016, Winter 2017, Spring 2018, Spring-Fall 2019, and Fall 2020 - Winter 2021. We can also see the end of an 11th event in Summer 2009. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_8.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_8.pdf}
\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

Figures 7 and 8 show the same comparison between the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively. For the 7th level, we see the same events as for the 8th level, both for the GPS data and the tremor count data. The wavelet decomposition is more noisy for the GPS data in the earliest part of the time series, between 2010 and 2013, but it does not seem that there are more slow slip events visible in the 7th level. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_7.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_7.pdf}
\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

For the 6th level detail, we see an additional event in Fall 2009 that is present both in the GPS and the tremor data. There are three small signals in the GPS data in Spring 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and are probably false detections. To summarize, all the 11 events present on the 7th and 8th level details of the wavelet decomposition are true detections, 12 of the 15 events present on the 6th level detail of the wavelet decomposition are true detections. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_6.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_6.pdf}
\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

@@

In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal. As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (levels 4 and 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog. Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events. Table 1 summarizes for each event its number as indicated in the catalog from ~\citet{MIC_2019}, the beginning and end times as indicated in the catalog from ~\citet{MIC_2019}, whether it is visible at the level 4 of the wavelet decomposition, whether it is visible at the level 5 of the wavelet decomposition, and whether it is part of a bigger magnitude 6 event. All 10 events that are sub-event of a bigger event are visible at both levels 4 and 5. However, this may be due because the bigger event is in at levels 6 to 8, and also at smaller time scales. For the 3 small events that are not part of a bigger event, only one is visible at both time scales, the other two are visible either for level either for level 5 of the wavelet decomposition. Therefore, it is difficult to conclude whether the method can indeed detect events of magnitude 5.

 \begin{table}
 \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
 \centering
 \begin{tabular}{c c c c c}
 \hline
 Event number & Time & Visible at level 4 & Visible at level 5 & Sub-event of bigger event \\
 \hline
 25 & 2010.62-2010.67 & Yes & Yes & Yes \\
 29 & 2011.42-2011.45 & No & Yes & No \\
 31 & 2011.62-2011.68 & Yes & Yes & Yes \\
 32 & 2011.65-2011.68 & Yes & Yes & Yes \\
 35 & 2012.66-2012.72 & Yes & Yes & Yes \\
 42 & 2013.70-2013.78 & Yes & Yes & Yes \\
 44 & 2014.12-2014-20 & Yes & No & No \\
 45 & 2014.40-2014.48 & Yes & Yes & No \\
 49 & 2014.66-2014.71 & Yes & Yes & Yes \\
 52 & 2014.91-2014.95 & Yes & Yes & Yes \\
 57 & 2015.98-2016.08 & Yes & Yes & Yes \\
 60 & 2017.11-2017.15 & Yes & Yes & Yes \\
 61 & 2017.20-2017.24 & Yes & Yes & Yes \\
 \hline
 \end{tabular}
 \end{table}

@@

@@

\acknowledgments
Enter acknowledgments, including your data availability statement, here.

>>>>

\title{Detection of slow slip events using wavelet analysis of GNSS recordings}

@@

\authors{A. Ducellier \affil{1}, K. C. Creager \affil{1}, D. A. Schmidt \affil{1}}

@@

\affiliation{1}{University of Washington}
%(repeat as many times as is necessary)

@@

\correspondingauthor{Ariane Ducellier}{ducela@uw.edu}

@@

\begin{abstract}
Slow slip events were discovered in many subduction zones during the last two decades thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks. Slow slip can last from a few days to several years and has a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles. In many places, tectonic tremor is also observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in GNSS noise. However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip. Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next. In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform. In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assume that there is a slow slip event whenever there is a peak in the wavelet signal. We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data. The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citep{MIC_2019}).
\end{abstract}

@@

In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data. For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor. For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded. They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8. They also observed many smaller bursts of tremor of duration 1 to 50 hours in between the big ETS events, without any detectable signal in the GPS data. However, based on the relationship between slow slip moment and number of hours of tremor for bigger events, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with the tremor bursts. This leads to a power-law relationship between seismic moment and number of events with a $b$-value close to one, similar to the distribution of normal earthquakes (\textcolor{red}{reference?}). \\

~\citet{FRA_2016} divided GPS time series observations into two groups: the first group contains days when slow seismicity (tectonic tremor and LFEs) is detected, the second group contains days when the numbers of tremor or LFEs is lower than a threshold. He then stacked separately the two groups of observations and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion. He also observed a cumulative displacement in the southern direction (for Guerrero) and the western direction (for Cascadia) corresponding to the release period when tremor and LFEs are observed. He was thus able to observe a reverse displacement corresponding to smaller slow slip events not directly observable in the GPS data. \\

However, in other subduction zones such as New Zealand, there is no clear relationship between tremor and slow slip occurrence and these methods cannot be applied to detect smaller slow slip events that produce a GNSS signal with an amplitude too small compared to the noise. We thus need other methods to be able to better detect and quantify slow slip. \\

Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next. Wavelet methods have been widely used for geophysical applications ~\citep{KUM_1997}. However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events ~\citep{SZE_2008,OHT_2010,WEI_2012,ALB_2019}. \\

~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005. They modeled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, Heaviside step functions corresponding to earthquakes and hardware upgrades, and a residual signal. They then applied a Gaussian wavelet transform to the residual time series to get the exact timing of the slow slip at each GPS station. The idea is that the wavelet transform allows us to analyze the signal both in the time and the frequency domains. A sharp change in the signal will be localized and seen at all levels of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform. \\

Instead of using wavelets in the time domain, ~\citet{OHT_2010} used 2D wavelet functions in the spatial domain to detect slow slip events. They designed the Network Stain Filter (NSF) to detect transient deformation signals from large-scale geodetic arrays. They modeled the position of the GPS station by the sum of the secular velocity, a spatially coherent field, site-specific noise, reference frame errors, and observation errors. The spatial displacement field is modeled by the sum of basis wavelets with time-varying weights.  Their method has been successfully used to detect a transient event in the Boso peninsula, Japan, and a slow slip event in the Alaska subduction zone ~\citep{WEI_2012}. \\

Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011. Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the ETS events. The noise component is assumed to be coherent between all tidal gauges, while the uplift signal is different provided that the gauges are far enough from each other. By stacking the tidal records, the uplift signals cancel each other while the noise signal is amplified. By stacking the details of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal. The authors were then able to clearly see a difference in uplift between the two tidal gauges at Port Angeles and Port Townsend. \\

In our study, we use a similar approach with a different reasoning. We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered. We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking. Finally, we suppose that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events. We use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia. Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may be currently undetected with standard methods. \\

@@

We focused our study on northwest Washington State. For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University. These are network solutions in ITRF2008 with phase ambiguities resolved. Solutions are computed with JPL/NASA orbits and satellite clocks. North, East, and Vertical directions are available. However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio. The wavelet method works best with data with zero mean, and no sharp discontinuities, so we use the cleaned dataset, that is GPS times series with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}. For the seismic data, we used the tremor catalog from \textcolor{red}{reference for the catalog. The following is to be modified. the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}. Tremor were detected and located using waveform envelope correlation and clustering and a centroid location is available for every given five-minute time window when tremor was detected. As the catalog starts in August 2009, we only looked at GPS data recorded in 2009 or later.} \\

@@

The MODWT of a time series can be defined for any length $N$. The time step of the wavelet vectors $\widetilde{W}_j$ and the scaling vector $\widetilde{V}_J$ is equal to the time step of the original time series. When we circularly shift the time series, the corresponding wavelet vectors, scaling vector, details and smooths are shifted by the same amount. The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series. The wavelet methods for time series analysis are explained in a more detailed way in ~\citep{PER_2000}). \\

@@

To increase the signal-to-noise ratio and be able to better detect slow slip events, we stack the signal over several GPS stations. We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal. We choose 16 points located on the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4). Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations. We thus have a stacked detail for each level 1 to 10 of the wavelet decomposition. \\

@@

We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. The result is shown in the top panel of Figure 6, where each line represents one of the locations. To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold. To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6. We used the opposite of the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks. In the tremor catalog from \textcolor{red}{reference?}, there are 17 tremor events with more than 150 hours of tremor recorded. The events are summarized in Table 1. The time of the event is the start date plus half the duration of the event. Although the latitudinal extension of the events is not always the same for the GPS data and for the tremor data, we identify the same 13 events in both 8th wavelet decompositions for the 8th level: January 2007, May 2008, May 2009, August 2010, August 2011, September 2012, September 2013, August-November 2014, January 2016, March 2017, June 2018, March-November 2019, and October 2020-January 2021. Although there are two events in the tremor catalog in August 2014 and November 2014, these two events are not distinguishable in the 8th level details and look more like a single event slowly propagating from South to North. The same phenomenon is observed in 2019 when two tremor events in March and November 2019 are merged into a single event propagating slowly from South to North. In 2020-2021, the wavelet decomposition of the tremor shows one event in the south in October-November 2020 and one event in the North in January 2021, but in the wavelet decomposition of the GPS data, these three events look like a single event propagating slowly from South to North. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.pdf}
\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

Figures 7 and 8 show the same comparison between the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively. For the 7th level, we see the same events as for the 8th level, both for the GPS data and the tremor count data. The wavelet decomposition is more noisy for the GPS data between 2010 and 2012, but it does not seem that there are more slow slip events visible in the 7th level. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.pdf}
\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

For the 6th level detail, we see an additional event in the South in Fall 2009 that is present both in the GPS and the tremor data. It may correspond to the northern extent of a big ETS event occurring in Fall 2009 south of the study area (event 19 in the ~\citet{MIC_2019} catalog). There are three small signals in the GPS data in Spring 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and are probably false detections. To summarize, all the 13 events present on the 7th and 8th level details of the wavelet decomposition are true detections, 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.pdf}
\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

 \begin{table}
     \caption{Big Episodic Tremor and Slip events. The duration and the number of tremor are from the tremor catalog of \textcolor{red}{reference?}. The event number and the magnitude are from the slow slip catalog of ~\citet{MIC_2019}.}
 \centering
 \begin{tabular}{c c c c c}
 \hline
 Time & Duration (days) & Number of tremor (hours) & Event number & Magnitude \\
 \hline
 2007.06 & 28 & 398 & 3 & 6.68 \\
 2008.36 & 25 & 402 & 10 & 6.56 \\
 2009.35 & 24 & 248 & 16 & 6.49 \\
 2010.63 & 29 & 518 & 24 & 6.54 \\
 2011.60 & 37 & 479 & 30 & 6.47 \\
 2012.72 & 37 & 620 & 34 & 6.54 \\
 2013.71 & 27 & 423 & 41 & 6.58 \\
 2014.65 & 15 & 190 & 48 & 6.03 \\
 2014.89 & 38 & 385 & 51 & 6.40 \\
 2016.11 & 43 & 421 & 54 & 6.79 \\
 2017.23 & 19 & 279 & 59 & 6.61 \\
 2018.49 & 22 & 381 & & \\
 2019.23 & 34 & 195 & & \\
 2019.88 & 16 & 205 & & \\
 2020.79 & 26 & 193 & & \\
 2020.86 & 12 & 162 & & \\
 2021.09 & 14 & 230 & & \\
 \hline
 \end{tabular}
 \end{table}

@@

In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal. As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (level 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog. Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events. Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event. \\

 \begin{table}
 \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
 \centering
 \begin{tabular}{c c c c}
 \hline
 Time & Event number & Magnitude & Sub-event of bigger event \\
 \hline
 2007.06 & 1 & 5.64 & Yes \\
 2007.08 & 2 & 5.91 & Yes \\
 2008.38 & 11 & 5.50 & Yes \\
 2009.16 & 14 & 5.50 & No \\
 2009.36 & 17 & 5.32 & Yes \\
 2010.63 & 25 & 5.76 & Yes \\
 2011.66 & 31 & 5.61 & Yes \\
 2011.66 & 32 & 5.32 & Yes \\
 2012.69 & 35 & 5.56 & Yes \\
 2013.74 & 42 & 5.71 & Yes \\
 2014.69 & 49 & 5.31 & Yes \\
 2014.93 & 52 & 5.39 & Yes \\
 2016.03 & 57 & 5.80 & Yes \\
 2017.13 & 60 & 5.43 & Yes \\
 2017.22 & 61 & 5.37 & Yes \\
 \hline
 \end{tabular}
 \end{table}

Figure 9 shows the 5th level detail wavelet decomposition of the GPS data. Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_michel_detail_5.pdf}
\caption{Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.}
\label{pngfiguresample}
\end{figure}

All 14 events that are sub-event of a bigger event are visible at level 5. However, this may be due because the bigger event are clearly seen at levels 6 to 8, and also at smaller time scales. The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition. However, some other events that are not in ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016. Therefore, it is difficult to make the difference between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.

@@

In this paper, we have used wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We used detrended GNSS data, applied the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also computed the cumulative number of tremor in the vicinity of the GNSS stations, detrended this signal, and applied the MODWT transform. In both time series, we could then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assumed that there is a slow slip event whenever there is a peak in the wavelet signal. We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data. The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citep{MIC_2019}). We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to make the difference between true detections and false detections.

@@

\acknowledgments
This work was funded by the grant from the National Science Foundation XXX. A.D. would like to thank Pr Donald Percival for introducing her to wavelet methods during his excellent class taught at University of Washington. The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA}. Figure 4 was done using GMT ~\citep{WES_1991}. The Python scripts used to analyze the data and make the  figures can be found on the  first author's Github account.
